### 2014-01-08 report.tex
added more sections

\documentclass[12pt, letterpaper]{article}

\title{Security Audit of Safeplug}
\author{Annie Edmundson and Anna Kornfeld Simpson}

\begin{document}

\maketitle

\section{Introduction}

\section{Background}

\section{Methodology}

\section{Results}

\section{Conclusion}

\input{bibliography}

\end{document}