### 2014-01-09 bibliography.tex

\begin{thebibliography}{99}

\bibitem{marvellhw} Marvell, ``88F6190 and 88F6192 Integrated Controller Hardware Specifications'' 2 December 2008. \url{http://www.marvell.com//embedded-processors/kirkwood/assets/HW_88F619x_OpenSource.pdf}

\bibitem{marvell} Marvell, ``Marvell 88F6192 SoC with Sheeva Technology'' 2009. \url{http://www.marvell.com//embedded-processors/kirkwood/assets/88F6192-003_ver1.pdf}

\bibitem{marvell2} Marvell, ``Marvell Alaska 88E1116R'' 2007. \url{http://www.marvell.com/transceivers/assets/Marvell-Alaska-88E1116R-Single-Port-GbE.pdf}

\bibitem{safeplug} Safeplug, \url{https://pogoplug.com/safeplug}.

\bibitem{wired} Safeplug makes it super-easy to harness Tor's anonymity at home, \url{http://www.wired.co.uk/news/archive/2013-11/22/safeplug-tor}.

\bibitem{bittech} Pogoplug launches Tor-powered Safeplug, \url{http://www.bit-tech.net/news/hardware/2013/11/25/pogoplug-safeplug/1}.

\bibitem{schneier} Schneier on Security: Tor Appliance.  \url{https://www.schneier.com/blog/archives/2013/11/tor_appliance.html}.

\bibitem{gigaom} Say hello to Safeplug, Pogoplug’s \$49 Tor-in-a-box for anonymous surfing, \url{http://gigaom.com/2013/11/21/say-hello-to-safeplug-pogoplugs-49-tor-in-a-box-for-anonymous-surfing/}.

\bibitem{techreview} Online Anonymity in a Box, for \$49, \url{http://www.technologyreview.com/news/521676/online-anonymity-in-a-box-for-49/}.

\bibitem{wireshark} Wireshark, \url{http://www.wireshark.org/}.

\end{thebibliography}