### 2014-01-13 goals.tex
modified the paper to include sections

\section{Goals}
\label{sec:goals}
TODO: Annie
In this project, we would like to do a security audit of the Safeplug and investigate the following questions.  How does the box pick which relay node on the Tor network to connect to?  What kind of authentication goes on between the box and the Tor node?  Does it use the same route of Tor relays every time?
Safeplug claims to offer in-box ad-blocking.  How does this work and does it remove any functionality?  Is there an easy way to whitelist things without turning Safeplug off entirely?

What are the configurations on the box?  Is it clear to users what information is being hidden and what information is being leaked?  The website talks a lot about IP addresses and warns users to clear cookies \cite{safeplug}, but is that sufficient to preserve anonymization across all web traffic (including Flash or Javascript supercookies that might not be cleared when cookies are cleared)?  Can user anonymity be leaked by a user’s behavior while still using the Safeplug?  What else must a user do in addition to using the Safeplug to further preserve anonymity?

Safeplug can act as a relay in the Tor network \cite{techreview}; how would widespread use of the box affect the Tor network?