### 2014-01-13 intro.tex

\section{Introduction}
\label{sec:intro}
    -Motivation
    -Tor
    -Safeplug
Safeplug is a product that launched last week from the cloud storage company Pogoplug, offering to let users browse the web with “complete security and anonymity” for the cheap price of \$49 \cite{safeplug}.  Safeplug offers Tor out of the box, with no additional software installation, by sitting between a user’s router and the internet \cite{wired}.  Pogoplug’s marketing pitch centers around the protection of user’s IP addresses by using Tor \cite{safeplug,bittech}.  Lots of folks are excited about the idea, but waiting to see a security audit, including Bruce Schneier \cite{schneier}.  Gigaom quotes Pogoplug CEO Dan Putterman as saying the box uses “vetted software” \cite{gigaom} but who knows what that means?