### 2014-01-13 procedure.tex

\section{Procedure}
\label{sec:procedure}
TODO: Anna
Safeplug is not open source, which makes a security audit much more challenging.  This leads to a debate on where a user will put his/her trust.  Despite not having the source code, we hope to get a Safeplug, set it up via their website, and use Wireshark to capture traffic and see what happens.  Wireshark allows us to capture and analyze network data \cite{wireshark}.  We are interested both in the communication between the clients and the Safeplug and between the Safeplug and the outside world.