### 2014-01-13 report.tex

\documentclass[12pt, letterpaper]{article}
\usepackage{url}

\title{Security Audit of Safeplug}
\author{Annie Edmundson and Anna Kornfeld Simpson}

\begin{document}

\maketitle

\input{intro}
\input{goals}
\input{procedure}
\input{results}
\input{related}
\input{conclusion}
\input{bibliography}

\end{document}