### 2014-01-13 results.tex

\section{Results}
\label{sec:results}
\subsection{User Experience}
TODO: Annie
    -Terms of Service
    -Setup to home router
    -Using the internet
    -Experience
    -Cookie experience (Fourthparty)

\subsection{Teardown}
TODO: Annie
    -Parts in the box
    -Possible attacks via USB/SD Card

\subsection{Traffic Analysis}
TODO: Anna
    -What's the udpate?
    -Phone home frequency
    -Does it do anything else?
    -How does the ad-blocking work?

\subsection{Gaining Access through SSH}
TODO: Anna
    -How we got access
    -What's the software
    -What/where are the settings
    -RPC server
        -Attacks from anyone on local network
        -Attack from a website?