### 2014-01-14 results.tex
split results in half

\section{Results}
\label{sec:results}
\subsection{User Experience}
\label{sec:ux}
TODO: Annie
    -Terms of Service
    -Setup to home router
    -Using the internet
    -Experience
    -Cookie experience (Fourthparty)

\subsection{Teardown}
\label{sec:tear}
TODO: Annie
    -Parts in the box
    -Possible attacks via USB/SD Card

\begin{figure}[htb]
\begin{center}
\includegraphics[width=\linewidth]{safeplug_top}
\caption{This is a figure.}
\end{center}
\end{figure}

\begin{figure}[htb]
\begin{center}
\includegraphics[width=\linewidth]{safeplug_bottom}
\caption{This is a figure.}
\end{center}
\end{figure}