### 2014-01-14 procedure.tex

\section{Procedure}
\label{sec:proc}
TODO: Anna
Safeplug is not open source, which makes a security audit much more challenging.  This leads to a debate on where a user will put his/her trust.  Despite not having the source code, we hope to get a Safeplug, set it up via their website, and use Wireshark to capture traffic and see what happens.  Wireshark allows us to capture and analyze network data \cite{wireshark}.  We are interested both in the communication between the clients and the Safeplug and between the Safeplug and the outside world.

We used the FourthParty plugin developed by Jonathan Mayer at Stanford to collect information about cookies and other browsing data during the client's use of the Safeplug \cite{fourthparty}.