### 2014-01-14 report.tex

\documentclass[12pt, letterpaper]{article}
\usepackage[hyphens]{url}
\usepackage{breakurl}
\usepackage[pdftex]{graphicx}
\usepackage[left=1in,top=1in,right=1in,bottom=1in,head=.5in,foot=.5in]{geometry}
\usepackage{setspace}

%\usepackage{caption}
\usepackage{subcaption}
%\usepackage[small,compact]{titlesec}
\usepackage{listings}
\lstnewenvironment{fileName}{%
\lstset{%
        basicstyle=\ttfamily,
        breaklines,
    }%
}{%
}

\title{Security Audit of Safeplug}
\author{Annie Edmundson and Anna Kornfeld Simpson}

\begin{document}

\maketitle

\input{intro}
\input{goals}
\input{procedure}
\input{results}
\input{resultsSoftware}
\input{related}
\input{conclusion}

\section{Acknowledgements}
We would like to express our gratitude to Josh Kroll for donating a lot of time and his computer to the routing effort.  We also thank whoever left the very nice set of screwdrivers on our desks!

This paper represents our own work in accordance with University regulations.

Anna Kornfeld Simpson and Annie Edmunson

\input{bibliography}

\end{document}