### 2014-01-14 related.tex

\section{Related Work}
\label{sec:related}

{\bf Tor.} There has been previous research analyzing Tor and evaluating Tor's security against realistic adversaries~\cite{tor2}.  Specifically, research has shown that it is possible to deanonymize users that use Tor.  Safeplug does not modify Tor, it simply uses it.  Despite warnings on the Tor project website that state no other configurations are guaranteed to make a user anonymous, except for the Tor Browser Bundle~\cite{torproject}.

{\bf Fingerprinting.}  Prior research has shown privacy attacks based on fingerprinting physical devices as well as through traffic analysis~\cite{fingerprint1, fingerprint2}.  We used Panopticlick to determine how unique our browser fingerprint is, and it was determined to be unique, despite Safeplug's anonymity guarantees~\cite{pano}.

{\bf Third-party and Supercookies.} Cookies and web tracking have been popular research areas lately.  Some related work regards user privacy leaks based on cookies~\cite{commercial1, commercial2}.