### 2014-01-27 report/design.tex
reorganized paper and added sections according to diebold paper

\section{Design and Operation of Safeplug}
\label{sec:design}