### 2014-01-27 report/implementattacks.tex

\section{Implementing Attacks}
\label{sec:implementattacks}