### 2014-01-27 report/abstract.tex
added TODO items in paper

\section{Abstract}
\label{sec:abstract}

TODO: Annie