### 2014-01-27 report/design.tex

\section{Design and Operation of Safeplug}
\label{sec:design}

\subsection{Hardware}
TODO: Annie

\subsection{Software}