### 2014-01-27 report/implementattacks.tex

\section{Implementing Attacks}
\label{sec:implementattacks}

\subsection{Gaining Access through SSH}

\subsection{Insider Attack}

\subsection{CSRF Attack}

TODO: Annie