### 2014-01-27 report/usability.tex

\section{Usability}
\label{sec:usability}

TODO: Annie