### 2014-02-04 report/discussion.tex
ideas for discussion section

\section{Discussion}
\label{sec:discussion}

-does this device have a real market/purpose?

-did not test on mobile devices