### 2014-02-08 report/bibliography.tex
finished related work and added items to bibliography

\begin{thebibliography}{99}

\bibitem{spymall} Appelbaum et. al. ``Shopping for Spy Gear: Catalog Advertises 
NSA Toolbox'' \emph{Der Spiegel} 29 December 2013 \url{www.spiegel.de/internatio
nal/world/catalog-reveals-nsa-has-back-doors-for-numerous-devices-a-940994.html}

\bibitem{nsa} Arther, Charles.  ``NSA scandal: what data is being monitored and how does it work?'' \url{http://www.theguardian.com/world/2013/jun/07/nsa-prism-records-surveillance-questions}.

\bibitem{fingerprint1} Cai, Xiang, et al. ``Touching from a distance: Website fingerprinting attacks and defenses.'' \emph{Proceedings of the 2012 ACM conference on Computer and Communications Security (CCS)}, 2012.

\bibitem{ceadmin} Colleton, Lee. ``Fwd: SSH on Safeplug'' \emph{Tor-talk mailing list} 2 January 2014. \url{http://archives.seul.org/or/talk/Jan-2014/msg00003.html}

\bibitem{tor} Dingledine, Roger, et. al. ``Tor: The second-generation onion router.'' \emph{USENIX Security Symposium}, 2004.

\bibitem{dyer} Dyer, Kevin P., Coull, Scott E., Ristenpart, Thomas, and Shrimpton, Thomas.  ``Peek-a-boo, i still see you: Why efficient traffic analysis countermeasures fail.'' \emph{In Proceedings of the 33rd Annual IEEE Symposium on Security and Privacy}, 2012.

\bibitem{feamster} Feamster, Nick, and Dingledine, Roger.  ``Location Diversity in Anonymity Networks.''  \emph{In ACM Workshop on Privacy in the Electronic Society (WPES)}, 2004.

\bibitem{fredrikson} Fredrikson, M., and Livshits, B. ``Repriv: Re-envisioning in-browser privacy.'' \emph{In Proceedings of the 2011 IEEE Symposium on Security and Privacy}, May 2011.

\bibitem{guha} Guha, S., Cheng, B., and Francis, P. ``Privad: Practical privacy in online advertising.'' \emph{In Proceedings of the 2011 USENIX Symposium on Networked Systems Design and Implementation}, April 2011.

\bibitem{bittech} Halfacree, Gareth. ``Pogoplug launches Tor-powered Safeplug'' \emph{bit-tech.net} 25 November 2013, \url{http://www.bit-tech.net/news/hardware/2013/11/25/pogoplug-safeplug/1}.

\bibitem{herrmann} Herrmann, Dominik, Wendolsky, Rolf, and Federrath, Hannes.  ``Website fingerprinting: attacking popular privacy enhancing technologies with the multinomial naive-bayes classifier.'' \emph{In Proceedings of the 2009 ACM workshop on Cloud computing security}.

\bibitem{tor2} Johnson, Aaron, et al. ``Users get routed: Traffic correlation on Tor by realistic adversaries.'' \emph{Proceedings of the 2013 ACM SIGSAC conference on Computer \& Communications Security (CCS)}, 2013.

\bibitem{dropbear} Johnston, Matt. ``Dropbear SSH.'' \url{https://matt.ucc.asn.au/dropbear/dropbear.html}

\bibitem{fingerprint2} Kohno, Tadayoshi et. al. "Remote physical device fingerprinting." Dependable and Secure Computing, IEEE Transactions on 2.2 (2005): 93-108.

\bibitem{lighttpd} Lighttpd. \url{http://www.lighttpd.net}

\bibitem{marvellhw} Marvell, ``88F6190 and 88F6192 Integrated Controller Hardware Specifications'' 2 December 2008. \url{http://www.marvell.com//embedded-processors/kirkwood/assets/HW_88F619x_OpenSource.pdf}

\bibitem{marvell} Marvell, ``Marvell 88F6192 SoC with Sheeva Technology'' 2009. \url{http://www.marvell.com//embedded-processors/kirkwood/assets/88F6192-003_ver1.pdf}

\bibitem{marvell2} Marvell, ``Marvell Alaska 88E1116R'' 2007. \url{http://www.marvell.com/transceivers/assets/Marvell-Alaska-88E1116R-Single-Port-GbE.pdf}

\bibitem{fourthparty} Mayer, Jonathan. ``FourthParty'' \url{http://fourthparty.info}

\bibitem{commercial2} Mayer, Jonathan R., and John C. Mitchell. "Third-party web tracking: Policy and technology." \emph{IEEE Symposium on Security and Privacy (SP)}, 2012.

\bibitem{gigaom} Meyer, David. ``Say hello to Safeplug, Pogoplug’s \$49 Tor-in-a-box for anonymous surfing'', \emph{GigaOm}, 21 November 2013. \url{http://gigaom.com/2013/11/21/say-hello-to-safeplug-pogoplugs-49-tor-in-a-box-for-anonymous-surfing/}.

\bibitem{murdoch} Murdoch, S.J., Danezis, G.  ``Low-Cost Traffic Analysis of Tor.'' \emph{In IEEE Symposium on Security and Privacy (Oakland)}, 2005.

\bibitem{murdoch2} Murdoch, S.J. and Zielinski, P.  ``Sampled Traffic Analysis by Internet-Exchange-Level Adversaries.'' \emph{In Privacy Enhancing Technologies (PET)}, 2007.

\bibitem{overlier} Overlier, L. and Syverson, P.  ``Locating Hidden Servers.'' \emph{In IEEE Symposium on Security and Privacy (Oakland)}, 2006.

\bibitem{panchenko} Panchenko, Andriy, Niessen, Lukas, Zinnen, Andreas, and Engel, Thomas.  ``Website fingerprinting in onion routing based anonymization networks.'' \emph{In Proceedings of the 10th Workshop on Privacy in the Electronic Society}, 2011.

\bibitem{pano} Panopticlick. \url{https://panopticlick.eff.org/}.

\bibitem{safeplug} Pogoplug. ``Safeplug'', \url{https://pogoplug.com/safeplug}.

\bibitem{privoxy} Privoxy. \url{https://www.privoxy.org}

\bibitem{schneier} Schneier, Bruce. ``Tor Appliance.'' \emph{Schneier on Security}, 27 November 2013. \url{https://www.schneier.com/blog/archives/2013/11/tor_appliance.html}.

\bibitem{techreview} Simonite, Tom. ``Online Anonymity in a Box, for \$49'', \emph{MIT Technology Review}, 21 November 2013. \url{http://www.technologyreview.com/news/521676/online-anonymity-in-a-box-for-49/}.

\bibitem{toubiana} Toubiana, V., Narayanan, A., Boneh, D., Nissenbaum, H., and Barocas, S. ``Adnostic: Privacy preserving targeted advertising.'' \emph{In Proceedings of the 2010 Network and Distributed System Security Symposium}, March 2010.

\bibitem{wired} Solon, Olivia. ``Safeplug makes it super-easy to harness Tor\'s anonymity at home,'' \emph{Wired UK}, 22 November 2013. \url{http://www.wired.co.uk/news/archive/2013-11/22/safeplug-tor}.

\bibitem{tormailinglist} Tor Mailing List.  \url{https://lists.torproject.org/pipermail/tor-talk/2013-November/031199.html}.

\bibitem{orbot} ``Tor on Android.'' \emph{Tor Project.} Accessed February 2014. \url{https://torproject.org/docs/android.html.en}

\bibitem{torproject} The Tor Project.  \url{https://www.torproject.org/}.

\bibitem{amorbot} The Tor Project. ``Orbot: Proxy with Tor.'' \emph{Google Android Market}. Accessed 7 February 2014, \url{https://play.google.com/store/apps/details?id=org.torproject.android}

\bibitem{wireshark} Wireshark, \url{http://www.wireshark.org/}.

\end{thebibliography}