### 2014-02-12 report/abstract.tex
rearranged sections

\begin{abstract}
We write this last.
\end{abstract}