### 2014-02-17 report/abstract.tex
First draft of an abstract

\begin{abstract}
We present the first public third-party security audit of Pogoplug's Safeplug device, which marketed ``complete security and anonymity online'' by using Tor technology to protect users' IP addresses.  We examine the hardware, software, and network behavior of the Safeplug device, as well as the user experience in comparison to other forms of web browsing.  Although the Safeplug appears to use Tor as advertised, lack of user education could easily allow users to be identified in ways that they might not expect.  Furthermore, an engineering vulnerability in the Safeplug's settings commands would allow an adversary internal or external to user's home network to silently disable Tor or modify other Safeplug settings, which completely overrules the security claims of the device.  Even if the engineering problem is fixed, the user experience challenges of this type of device make it inferior to the existing standard in anonymity methods, particularly the Tor Browser Bundle.
\end{abstract}